---
title: "End term Hackathon - Patterns/Rules Generation"
author: "Vigneshwar Srinivasan "
output:
  html_document:
    toc: yes
    toc_depth: '3'
---



#Clear working space
```{r}

rm(list = ls(all=TRUE))

```

# Agenda 

* Get the data

* Data Pre-processing

* Build a model

* Predictions

* Communication

# Reading & Understanding the Data

## Data Domain and Format

## Read the Data with loading of all packages

Ensure the dataset is located in current working directory, else you can change your working directory using the "setwd()" function.
Loading all reqired working packages

```{r}
library(lattice)
library(ggplot2)
library(caret)
library(grid)
library(DMwR)
library(stats)
library(arm)
library(MASS)
library(car)
library(corrplot)
library(ROCR)
library(gplots)
library(corrplot)

getwd()

accident_data1 <-read.csv(file = "Train.csv")
collision_data <-read.csv("NumberOfvehiclesbyCollosion.csv")
base_data <-read.csv("Base_Vehicle_Data.csv")

```

## Understand the data

* Data aggregation and recast

```{r}
library(reshape2)
new_data <-base_data[,c(1,2,3)]
head(new_data)
require(reshape2)
new_base_data <-dcast(new_data,Collision.Reference.No. ~ Vehicle.Type, value.var= "Vehicle.Reference.No", fun.aggregate = length)
head(new_base_data)

```

#Merging all datasets

```{r}
head(base_data)
head(new_base_data)
head(accident_data1)
str(accident_data1)
head(collision_data)

base_data_altered <- base_data[,!colnames(base_data) %in% c("Vehicle.Type","Vehicle.Reference.No")]
head(base_data_altered)

accident_data_1 <-merge(new_base_data, accident_data, by = "Collision.Reference.No.")
accident_data_2 <-merge(accident_data_1, collision_data, by = "Collision.Reference.No.")
accident_data_final <-merge(accident_data_2, base_data_altered, by = "Collision.Reference.No.")
dim(accident_data_final)

```


## Data Description

Collision Severity - Target variable 

## Understand the data

* Use the str() function to get the dimensions and types of attributes in the dataset

```{r}

str(accident_data_final)

View(accident_data_final)

```


* Use the summary() function to understand the distribution of variables in the dataset

```{r}

summary(accident_data_final)

```

* Use the head() and tail() functions to get a look at the data

```{r}

head(accident_data_final)

```


# Data Pre-processing

## Missing Values

* Check the number of missing values in the data frame and correlation plot between the independent variables

```{r}

colSums(is.na(accident_data_final))

table(colSums(is.na(accident_data_final)))

```

##Data pre-process

* Factor conversion

```{r}

accident_data_final$Collision.Severity <- as.factor(as.character(accident_data_final$Collision.Severity))
accident_data_final$Weekday.of.Collision <- as.factor(as.character(accident_data_final$Weekday.of.Collision))
accident_data_final$Day.of.Collision <- as.factor(as.character(accident_data_final$Day.of.Collision))
accident_data_final$Month.of.Collision <- as.factor(as.character(accident_data_final$Month.of.Collision))
accident_data_final$Hour.of.Collision..24.hour. <- as.factor(as.character(accident_data_final$Hour.of.Collision..24.hour.))
accident_data_final$Carriageway.Type <- as.factor(as.character(accident_data_final$Carriageway.Type))
accident_data_final$Junction.Detail <- as.factor(as.character(accident_data_final$Junction.Detail))
accident_data_final$Junction.Control <- as.factor(as.character(accident_data_final$Junction.Control))
accident_data_final$Pedestrian.Crossing....Human.Control <- as.factor(as.character(accident_data_final$Pedestrian.Crossing....Human.Control))
accident_data_final$Pedestrian.Crossing....Physical.Control <- as.factor(as.character(accident_data_final$Pedestrian.Crossing....Physical.Control))

accident_data_final$Light.Conditions <- as.factor(as.character(accident_data_final$Light.Conditions))
accident_data_final$Weather.Conditions <- as.factor(as.character(accident_data_final$Weather.Conditions))
accident_data_final$Road.Surface.Conditions <- as.factor(as.character(accident_data_final$Road.Surface.Conditions))
accident_data_final$Special.Conditions.at.Site <- as.factor(as.character(accident_data_final$Special.Conditions.at.Site))
accident_data_final$Number.of.Vehicles <- as.factor(as.character(accident_data_final$Number.of.Vehicles))
accident_data_final$Towing.and.Articulation <- as.factor(as.character(accident_data_final$Towing.and.Articulation))
accident_data_final$Vehicle.Manoeuvre <- as.factor(as.character(accident_data_final$Vehicle.Manoeuvre))
accident_data_final$Junction.Location.of.Vehicle.at.Time.of.Impact <- as.factor(as.character(accident_data_final$Junction.Location.of.Vehicle.at.Time.of.Impact))
accident_data_final$Skidding...Overturning <- as.factor(as.character(accident_data_final$Skidding...Overturning))
accident_data_final$First.Object.Hit.in.Carriageway <- as.factor(as.character(accident_data_final$First.Object.Hit.in.Carriageway))
accident_data_final$Vehicle.Leaving.Carriageway <- as.factor(as.character(accident_data_final$Vehicle.Leaving.Carriageway))
accident_data_final$First.Object.Hit.off.Carriageway <- as.factor(as.character(accident_data_final$First.Object.Hit.off.Carriageway))
accident_data_final$First.Point.of.Impact <- as.factor(as.character(accident_data_final$First.Point.of.Impact))
accident_data_final$Sex.of.Driver <- as.factor(as.character(accident_data_final$Sex.of.Driver))
accident_data_final$Hit.and.Run <- as.factor(as.character(accident_data_final$Hit.and.Run))
accident_data_final$Foreign.Registered.Vehicle <- as.factor(as.character(accident_data_final$Foreign.Registered.Vehicle))
accident_data_final$Towing.and.Articulation <- as.factor(as.character(accident_data_final$Towing.and.Articulation))

str(accident_data_final)

```


#New Data Structure
```{r}

colSums(is.na(accident_data_final))

library("imputeMissings")
accident_data_fin1 <-compute(accident_data_final, method = "median/mode" )
accident_data_fin <-impute(accident_data_final, object = NULL, method = "median/mode", flag = FALSE )

colSums(is.na(accident_data_final))

```


## 

* Scaling the  data

```{r}

library(lattice)
library(ggplot2)
library(caret)
std_model <-preProcess(accident_data_fin[, !names(accident_data_fin) %in% c("Collision.Severity")], method = c("center", "scale"))
accident_data_fin[,!names(accident_data_fin) %in% c("Collision.Severity")] <- predict(object = std_model, newdata = accident_data_fin[, !names(accident_data_fin) %in% c("Collision.Severity")])

accident_data_fin <- accident_data_fin[,!colnames(accident_data_fin) %in% c("Collision.Reference.No")]

```
#Generatign the rules only for the Collision.Severity (severity case) alone as of now

# Decision Tree Building


### Data transformation with Train/Test Split


```{r}
set.seed(786)
library(lattice)
library(ggplot2)
library(caret)
train_rows <-createDataPartition(accident_data_fin$Collision.Severity, p = 0.7, list = FALSE)
train_data <-accident_data_fin[train_rows,]

test_data <-accident_data_fin[-train_rows,]


```

##Decision Tree

```{r}
library(C50)
c5_tree <- C5.0(Collision.Severity ~ . , train_data)

c5_rules <- C5.0(Collision.Severity ~ . , train_data, rules = T)

```

## Variable importance in trees

```{r}
C5imp(c5_tree, metric = "usage")
```


#Rules from trees
```{r}
summary(c5_rules)

```


## Prediction on test data

```{r}
preds <- predict(c5_tree, test_data)

```


## Model accuracy


```{r}
library(caret)

confusionMatrix(preds, test_data$Collision.Severity)

```




